---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kutils

<!-- badges: start -->
[![License: AGPL v3](https://img.shields.io/badge/License-AGPLv3-blue.svg)](https://www.gnu.org/licenses/agpl-3.0)
[![Experimental](https://img.shields.io/badge/status-experimental-orange.svg)](https://github.com/kbosirany/kutils)
<!-- badges: end -->

## Overview

kutils provides utility functions for data management and configuration handling in R projects. The package wraps and extends functionality from the [fairify](https://forgemia.inra.fr/umr-g-eau/fairify.git) package to streamline common data operations.

## Features

- **Configuration Management**: Load and manage project configurations with `loadConfig()`
- **Data Path Handling**: Resolve and cache data paths with `getDataPath()`
- **Cache Management**: Clean cached data with `cleanDataCache()`

## Installation

You can install the development version of kutils from GitHub:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("kbosirany/kutils")
```

## Usage

### Configuration Management

Load configuration from a YAML file:

```{r eval=FALSE}
library(kutils)

# Load default configuration
cfg <- loadConfig()

# Load custom configuration file
cfg <- loadConfig(userFile = "my_config.yml")
```

### Data Path Handling

Resolve data paths with caching support:

```{r eval=FALSE}
# Get data path with caching enabled (default)
data_path <- getDataPath("my_data.csv")

# Disable caching
data_path <- getDataPath("my_data.csv", use_cache = FALSE)

# Use custom configuration
cfg <- loadConfig("my_config.yml")
data_path <- getDataPath("my_data.csv", cfg = cfg)
```

### Cache Management

Clean the data cache when needed:

```{r eval=FALSE}
# Clean default cache
cleanDataCache()

# Clean custom cache location
cleanDataCache(cache = "/path/to/cache")
```

## Configuration File

kutils uses YAML configuration files for managing data paths and modes. The default configuration is located at `inst/config.yml`. You can create your own configuration file following this structure:

```yaml
data:
  mode: local  # or 'remote', 'production', etc.
  local: /path/to/local/data
  remote: /path/to/remote/data
```

## Dependencies

- R (>= 4.0.0)
- fairify

## Development

### Running Tests

```{r eval=FALSE}
# Run all tests
devtools::test()

# Run specific test file
devtools::test_file("tests/testthat/test-test.R")
```

### Building Documentation

```{r eval=FALSE}
# Generate documentation
devtools::document()

# Build pkgdown site
pkgdown::build_site()
```

## Contributing

Contributions are welcome! Please feel free to submit a Pull Request.

## License

This project is licensed under the AGPL-3.0 or later - see the [LICENSE](LICENSE) file for details.

## Author

Kevin Bosirany Orlando  
Email: kevinbosirany@gmail.com  
ORCID: [0009-0009-2784-3108](https://orcid.org/0009-0009-2784-3108)

## Acknowledgments

This package builds upon the [fairify](https://forgemia.inra.fr/umr-g-eau/fairify.git) package developed by UMR G-EAU.
